Supporting online material for\\ \emph{Effectiveness of UNAIDS targets and HIV vaccination across 127 countries}

### 2016-08-15 Documents/supplementary_text.tex
Added stubs for other 3 data tables.

Model source code: \url{https://github.com/janmedlock/HIV-95-vaccine/}.

@@

Several sources provided information to parameterize our model at the
country level. Demographic data, including population growth rate
\cite{WorldBankpg}, death rate
\cite{World_Development_Indicators2013-ee}, and number of people aged
15--49 years \cite{The_World_Bank2016-fd} were obtained from the World
Bank. Longitudinal HIV incidence and prevalence estimates for the
15--49 age group, spanning from as early as 1990 to 2015, was
primarily derived from the AIDSinfo database produced by UNAIDS
\cite{Unaids2016-an}. Other sources included UNAIDS Country Progress
Reports \cite{Unaids2016-am} published from 2012 to 2016 and AIDS Data
Hub \cite{AIDSdatahub-fg}, which compiles data from UNAIDS, UNICEF,
WHO, and the Asian Development Bank. These sources were also used to
inform the initial conditions for the model: the number of people who
have been diagnosed with HIV, who are on ART, and who have viral
suppression or have been retained on treatment for at least 12
months. Where data were not available from these sources, alternative
resources such as peer-reviewed journal articles and country health
ministry reports were consulted.  We found sufficient data to
parametrize our mathematical modeling for 127 countries.  See
\autoref{data_source_treatment_cascade}--\ref{data_source_regions} for
full information on data sources for each country.

\begin{table}
  \centering
  \tiny
  \input{treatment_cascade.tex}
  \caption{Data sources on the treatment cascade.}
  \label{data_source_treatment_cascade}
\end{table}

@@

\begin{table}
\begin{center}
  \begin{tabular}{cp{4cm}llc}
\hline
    Symbol & Definition & Value & Sampling Distribution & Source \\ \hline
    $\delta$	& Rate of leaving acute phase of HIV infection & 4.1379/year & T (4.1379, 2, 9.6) & \cite{Hollingsworth2008-iy} \\
    $\sigma$	& Rate of developing AIDS among non-virally suppressed individuals
                       & 0.1064/year & & \cite{Morgan2002-cq} \\
    $\gamma$ & Rate of viral suppression & 1/year & U (0.5, 1.5) & \cite{Currie2009-yz} \\
    $\upsilon$	& Death rate from AIDS & 0.5/year &  & \cite{Morgan2002-cq} \\
    $\omega$	& Reduction in life years for treated PLHIV & 5 & U (5, 8) & \cite{Unaids2014-ue, Samji2013-kf}\\
    $\beta_{A}$	& Transmission probability during acute phase & 0.0082 & T (0.0082, 0.0039, 0.0150) & \cite{Skarbinski2015-ni,Wawer2005-us}\\
    $\beta_{U}$	& Transmission probability after acute phase & 0.0014 & T (0.0014, 0.00077, 0.00251) & \cite{Hughes2012-so}\\
    $\varepsilon$	& Proportional reduction of transmission with treatment & 0.08 & $\beta$ (0.08, 0.002, 0.57)& \cite{Donnell2010-xo}\\
    $n$			& Sex acts per year & 102 & U (96, 108) & \cite{Wawer2005-us,Abdool_Karim2010-cm}\\ \hline
  \end{tabular}
\caption{Parameters used in model. (T: triangular distribution, U:
  uniform distribution, $\beta$: beta distribution).}
\end{center}
\label{model_param}
\end{table}

@@

The targets for the proportions $p_D$, $p_T$, and $p_V$ in the achievement of 90---90---90
all operate under the following assumption:
if the initial level $p_x(0)$ is less than $90\%$, the proportion will increase
linearly in the first 5 years and then stay at $90\%$ for the
remaining time; and if the initial level is at $90\%$ or above, the proportion
will remain constant for the duration of the study:
\begin{equation}
  p_x^*(t)
  = p_x(0) + \left[
    \max\left\{0.9, p_x(0)\right\} - p_x(0)
  \right]
  \min\left\{\frac{t}{5}, 1\right\},
\end{equation}
for $x \in \{D, T, V\}$.

@@

95---95---95 goal is implemented subsequently after meeting
90---90---90 by2020 by  increasing $p_D$, $p_T$, and $p_V$ to
95---95---95 levels by 2030 linearly. \comment{@Jan, We use the
  Heaveside function here too?}

>>>>

\section*{To do}

\begin{itemize}
\item Finish tables
\item Reformat transmission rate figure
\end{itemize}

We developed a continuous-time, compartmental model consisting of 8
health states: susceptible to HIV infection, vaccinated against HIV,
acute HIV infection, undiagnosed HIV infection, diagnosed but
untreated HIV infection, treated without achieving viral suppression,
achieved viral suppression, and having AIDS (Supplementary text). We
found sufficient country-level demographic data, as well as estimates
of longitudinal HIV incidence and prevalence for the 15–49 age group,
spanning from as early as 1990 to 2015 to parameterize our model for
127 countries. See Supporting text for information on data sources.

Using available historical data for incidence and prevalence, and
published transmission estimates (13), we estimated transmission rates
for people with acute infection, unsuppressed viral load and
suppressed viral load. Our calibration method provided a recent
snapshot of country-specific transmission, while using historical data
to smooth out variations. See Supporting text for detailed information
on estimating the transmission rate.

Projections were generated for each country through model simulations
using 1000 random samples from published parameter distributions
(table S1, Supplementary text) and the country-specific estimated
distribution of the transmission rate. We evaluated the global and
country-level impacts of meeting the UNAIDS targets and vaccine
deployment on four outcome measures: HIV incidence, PLHIV, PLAIDS, and
AIDS-related deaths between 2015 and 2035. Our base-case scenario
assumed maintenance of the status quo: that 2015 levels of the
proportion of PLHIV diagnosed (pD), proportion of those diagnosed on
ART (pT), and proportion of those on treatment who have achieved viral
suppression (pV) continue unchanged until 2035. The intervention
scenarios were (1.) the attainment of the 90–90–90 target by linear
increases in pD, pT, and pV until 2020 followed by maintenance of
those levels until 2035, and (2.) linear increase in the diagnosis and
treatment levels to the 90–90–90 target by 2020, a linear increase
from 90–90–90 to the 95–95–95 target by 2030, and maintenance of the
95–95–95 levels to 2035. (For 90–90–90, if any initial level was above
90\%, it was kept constant going forward so that the intervention did
not worsen coverage.  Likewise, for 95–95–95, any initial level above
95\% was kept constant.)  For vaccination, we assumed deployment of a
vaccine with 50\% efficacy in 2020 with coverage increasing linearly
from 0\% to 50\% at 2 years post-rollout, continuing on the same
linear trend to 70\% at 2.8 years post-rollout, and maintaining
constant 70\% coverage thereafter. We also performed a partial rank
correlation coefficient (PRCC) sensitivity analysis for each model
parameter and a scenario analysis to account for uncertainty in the
profile of the prospective vaccine. See Supplementary text for more
details.

\section{Source code}

The source code for the model and analysis tools is available at
\url{https://github.com/janmedlock/HIV-95-vaccine/}.

@@

Several sources provided information to parameterize our model at the
country level. Demographic data, including population growth rate
\cite{WorldBankpg}, death rate
\cite{World_Development_Indicators2013-ee}, and number of people aged
15--49 years \cite{The_World_Bank2016-fd} were obtained from the World
Bank. Longitudinal HIV incidence and prevalence estimates for the
15--49 age group, spanning from as early as 1990 to 2015, was
primarily derived from the AIDSinfo database produced by UNAIDS
\cite{Unaids2016-an}. Other sources included UNAIDS Country Progress
Reports \cite{Unaids2016-am} published from 2012 to 2016 and AIDS Data
Hub \cite{AIDSdatahub-fg}, which compiles data from UNAIDS, UNICEF,
WHO, and the Asian Development Bank. These sources were also used to
inform the initial conditions for the model: the number of people who
have been diagnosed with HIV, who are on ART, and who have viral
suppression or have been retained on treatment for at least 12
months. Where data were not available from these sources, alternative
resources such as peer-reviewed journal articles and country health
ministry reports were consulted.  We found sufficient data to
parametrize our mathematical modeling for 127 countries.  See
\hyperref[data_source_treatment_cascade]{Tables
  \ref*{data_source_treatment_cascade}}--\ref{data_source_regions} for
full information on data sources for each country.

\begin{table}
  \centering
  \tiny
  \input{treatment_cascade}
  \caption{Data used for the treatment cascade.}
  \label{data_source_treatment_cascade}
\end{table}

\begin{table}
  \centering
  \tiny
  % \input{prevalence}
  \caption{Data sources for prevalence.}
  \label{data_source_prevalence}
\end{table}

\begin{table}
  \centering
  \tiny
  % \input{incidence}
  \caption{Data sources for prevalence.}
  \label{data_source_incidence}
\end{table}

\begin{table}
  \centering
  \tiny
  % \input{regions}
  \caption{Countries by UNAIDS region.}
  \label{data_source_regions}
\end{table}

@@

\begin{table}
  \begin{center}
    \begin{tabular}{cp{4cm}llc}
      \hline
      Symbol & Definition & Value & Sampling Distribution & Source \\ \hline
      $\delta$	& Rate of leaving acute phase of HIV infection & 4.1379/year & T (4.1379, 2, 9.6) & \cite{Hollingsworth2008-iy} \\
      $\sigma$	& Rate of developing AIDS among non-virally suppressed individuals
                          & 0.1064/year & & \cite{Morgan2002-cq} \\
      $\gamma$ & Rate of viral suppression & 1/year & U (0.5, 1.5) & \cite{Currie2009-yz} \\
      $\upsilon$	& Death rate from AIDS & 0.5/year &  & \cite{Morgan2002-cq} \\
      $\omega$	& Reduction in life years for treated PLHIV & 5 & U (5, 8) & \cite{Unaids2014-ue, Samji2013-kf}\\
      $\beta_{A}$	& Transmission probability during acute phase & 0.0082 & T (0.0082, 0.0039, 0.0150) & \cite{Skarbinski2015-ni,Wawer2005-us}\\
      $\beta_{U}$	& Transmission probability after acute phase & 0.0014 & T (0.0014, 0.00077, 0.00251) & \cite{Hughes2012-so}\\
      $\varepsilon$	& Proportional reduction of transmission with treatment & 0.08 & $\beta$ (0.08, 0.002, 0.57)& \cite{Donnell2010-xo}\\
      $n$			& Sex acts per year & 102 & U (96, 108) & \cite{Wawer2005-us,Abdool_Karim2010-cm}\\ \hline
    \end{tabular}
    \caption{Parameters used in model. (T: triangular distribution, U:
      uniform distribution, $\beta$: beta distribution).}
    \label{model_param}
  \end{center}
\end{table}

@@

The targets for the proportions $p_D$, $p_T$, and $p_V$ in the achievement of 90--90--90
all operate under the following assumption:
if the initial level $p_x(0)$ is less than $90\%$, the proportion will increase
linearly in the first 5 years and then stay at $90\%$ for the
remaining time; and if the initial level is at $90\%$ or above, the proportion
will remain constant for the duration of the study:
\begin{equation}
  p_x^*(t)
  = p_x(0) + \left[
    \max\left\{0.9, p_x(0)\right\} - p_x(0)
  \right]
  \min\left\{\frac{t}{5}, 1\right\},
\end{equation}
for $x \in \{D, T, V\}$.

@@

95--95--95 goal is implemented subsequently after meeting 90--90--90
by 2020 by increasing $p_D$, $p_T$, and $p_V$ to 95--95--95 levels by
2030 linearly. \comment{@Jan, We use the Heaveside function here too?}
